Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries

### 2022-03-04 Manuscript/manuscript_text2.tex
Updated manuscript

Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth (Ramankutty \emph{et al.} 2018).
This has allowed increases in human population and increased global stability in production (Foley \emph{et al.} 2011).
However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ({``multifunctionality''} \emph{sensu} Mitchell \emph{et al.} 2014; Frei \emph{et al.} 2018; but see also Karp \emph{et al.} 2018).
Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.

@@

Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure 1, Scenario 2).
Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Garibaldi \emph{et al.} 2011) and wind speed is reduced directly downwind of treelines (de Jong \& Kowalchuk 1995), so positive effects will be concentrated close to the boundary.
Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).

Unfortunately, data to inform these decisions are often minimal, as observing boundary effects requires large amounts of expensive and time-consuming crop yield data.
This means that studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Brandle \emph{et al.} 1984; Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
Meta-analyses are a common solution to this problem (e.g. Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs (Steel 2017; Osorio \emph{et al.} 2018).
However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems (Redhead \emph{et al.} 2020; Hünicken \emph{et al.} 2021); typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.

In this paper we use a precision yield dataset from Alberta, Canada to ask the following questions:
1. How does crop yield change with distance from the edge of field?
2. Does this depend on type of field edge?
3. Is there an intermediate distance where yield is maximized or variance is minimized?
To our knowledge, only (Osorio \emph{et al.} 2018) has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.

@@

Precision yield data were collected directly from farmers across Alberta (Figure S1).
Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
We converted raw data to a standard csv format using Ag Leader SMS®.
85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
Individual fields ranged from 5.9 to 369 ha in area (mean: 72.2 ha), and contained between 1 and 6 years of data (mean: 3.1), for a total of 14.4 million data points.

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix B for R code):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Removed outliers and ``inliers'' using the spatial filtering procedure from (Vega \emph{et al.} 2019)
\item
  Removed unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), and values of zero or less
\item
  Removed extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
\item
  Removed points with large changes (\textgreater20\%) in ground speed
\item
  Removed extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between data collection
\item
  Sub-sampled filtered data to 50,000 data points per field (reduce computation time)
\end{enumerate}

Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data (Agriculture and Agri-Food Canada 2018).
Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions.
For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Yield = weight of crop gathered per area of harvest (T/ha)
\item
  Boundary Distance = distance from field boundary type \emph{i} (m)
\item
  Easting, Northing = distance from centre of field (m)
\item
  f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
\end{enumerate}

In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure 1).
All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.38, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.5, Wickham 2016; and 0.4.0, Kassambara 2020).

The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 80), we decided to leave the models as-is.
Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
That is, combining of fields occurs in patterns (e.g.~east-to-west) so we cannot tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
Non-standard combine driving patterns may be able to gather data in a way that can detect sensor drift, but this is beyond the scope of this research.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors.
\item
  Use posterior sample to simulate a new smoother from each field.
\item
  Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this ``meta-smoother.''
\item
  Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
  This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
  Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.
\end{enumerate}

@@

Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
Distances from boundaries were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).

Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure 4 top row).
Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
Yield variability (Figure 4 bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.

@@

Pea crops were less well-represented in our data (N = 39 fields), and therefore the effects of boundaries tended to be less consistent.
Average pea yields (Figure 6, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
Conversely, yield variability (Figure 6, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.

@@

Within individual fields, both the average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure 1).
There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.

@@

Field boundary characteristics, such as height or species composition, may play an important role in mediating spillover, but our definitions of boundaries were necessarily loose, as we used satellite imagery to manually classify them.
Therefore, it may be that other aspects of field boundaries are more important for spillover of ecosystems services into fields.
For example, larger or more functionally diverse (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011; Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction influences predatory beetle communities (Sutter \emph{et al.} 2018).
These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
Finally, the relevant spatial scale of boundary effects are often related to the height or windward orientation of the shelterbelt (Kort 1988; Kowalchuk \& de Jong 1995), and precision yield data has the potential to show this in far greater detail than previous studies.

@@

We thank Trent Clark, Dean Hubbard, Alvin French, and Kristina Polziehn for providing yield data for us, and providing insight on their yield data.
We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-AAFC_data}{}%
Agriculture and Agri-Food Canada. (2018). \emph{{Annual Crop Inventory}}. Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}. Last accessed 6 February 2019.

@@

@@

\leavevmode\hypertarget{ref-statscan_canola2014}{}%
Statistics Canada. (2014). Table 001-0010 - estimated areas, yield, production and average farm price of principal field crops, in metric units, annual.

>>>>

Intensive agricultural production has increased over the last 100 years -- both in terms of area and intensity, such as enhanced tillage or the addition of synthetic fertilizers and pesticides -- and agricultural land now makes up over a third of ice-free land on Earth (Ramankutty \emph{et al.} 2018).
This has allowed increases in human population and increased global stability in production (Foley \emph{et al.} 2011).
However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ({``multifunctionality''} \emph{sensu} Mitchell \emph{et al.} 2014; Frei \emph{et al.} 2018; but see also Karp \emph{et al.} 2018).
Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.

@@

Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, soil compaction, or patchy herbicide application leading to higher weed competition (Figure 1, Scenario 2).
Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Garibaldi \emph{et al.} 2011) and wind speed is reduced directly downwind of treelines (de Jong \& Kowalchuk 1995), so positive effects will be concentrated close to the boundary.
Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Pywell \emph{et al.} 2015; Van Vooren \emph{et al.} 2017).

Unfortunately, data to inform these decisions are often minimal, as observing boundary effects requires large amounts of expensive and time-consuming crop yield data.
This means that studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Brandle \emph{et al.} 1984; Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
Meta-analyses are a common solution to this problem (e.g. Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs (Steel 2017; Pywell \emph{et al.} 2015; Osorio \emph{et al.} 2018).
However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems (Redhead \emph{et al.} 2020; Hünicken \emph{et al.} 2021); typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.

Within the Canadian prairie provinces, Alberta has roughly 58300 km$^2$ of cultivated agriculture (excluding pasture), of which the majority is spring wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), barley (\emph{Hordeum vulgare}), and peas (\emph{Pisum sativum}) (36.6\%, 30.8\%, 17.1\%, and 7.0\% in 2020, respectively; Agriculture and Agri-Food Canada 2022).
Roughly 85\% of growers in western Canada report having precision yield monitoring capability (Steel 2017), but these data are seldom used to answer large-scale agroecological questions.
In this paper we use a precision yield dataset from Alberta to ask the following questions:
1. How does crop yield change with distance from the edge of field?
2. Does this depend on type of field edge?
3. Is there an intermediate distance where yield is maximized or variance is minimized?
To our knowledge, only Pywell \emph{et al.} (2015) and Osorio \emph{et al.} (2018) have studied the effects of field boundaries using precision yield data, making this study one of the first of its kind, and even more so within a Canadian context.

@@

Precision yield data were collected directly from farmers across Alberta.
Farmers were solicited for yield data through local agronomists, and we received data from 5 growers for 87 individual fields across 7 years (2014-2020), for a total of 298 field-years of data (Figure S1).
We converted raw data to a standard csv format using Ag Leader SMS®.
85\% of the crop types where either wheat, canola, or peas, three of the most common crops in rotation in Alberta (Agriculture and Agri-Food Canada 2022).
Crop rotations in Alberta usually consist of a combination of wheat - canola - pulses - barley, with other crops being included dependent on the regional growing conditions (e.g. sunflowers, flax, corn).
The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
Individual fields ranged from 5.9 to 369 ha in area (mean: 72.2 ha), and contained between 1 and 6 years of data (mean: 3.1), for a total of 14.4 million data points.

Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
Fields were all conventional (standard herbicide and fungicide application), and seeding and application rates were constant within fields, so we did not consider these inputs in our analysis.
Precision yield data can be highly variable and is prone to extreme outliers, especially when ground speeds are low, during turning on headlands, or when the combine is changing speed (Arslan \& Colvin 2002; Griffin \emph{et al.} 2007; Whelan \& Taylor 2013).
Recording data between different fields can also cause extreme ``jumps" in location, and we verified that all the data points were located within the given field boundaries.
Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix B for R code):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item Removed outliers and ``inliers'' using the spatial filtering procedure from Vega \emph{et al.} (2019)
\item Removed unrealistic high yield values (10.75 t/ha for wheat, 8 t/ha for canola and peas), and values $\leq$ zero 
\item Removed extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
\item Removed points with large changes (\textgreater20\%) in ground speed
\item Removed extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between data collection
\item Sub-sampled filtered data to 50,000 data points per field (reduce computation time)
\end{enumerate}

Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data.
Edge location and type were usually clearly visible from the Google Earth imagery (CNES/Airbus imagery, 1.5 m resolution), but if they were ambiguous we used classified Agriculture and Agri-Food Canada (2022, 30 m resolution) data to confirm the land cover types surrounding each field.
The location of a given crop edge is flexible and can change yearly, depending on planting and emergence conditions.
For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
This makes consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item Yield = weight of crop gathered per area of harvest (t/ha)
\item Boundary Distance = distance from field boundary type \emph{i} (m)
\item Easting, Northing = distance from centre of field (m)
\item f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
\end{enumerate}

In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g. Kriging, SPDE approximations) but this was computationally unfeasible with 50,000 data points per field.
Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that the functional form of boundary effects would not be as complex as within-field variation (as in Figure 1).
All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.38, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.5, Wickham 2016; and 0.4.0, Kassambara 2020).

The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e. extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 80), we decided to leave the models as-is.
Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
That is, combining of fields occurs in patterns (e.g. east-to-west) so we cannot tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
Non-standard combine driving patterns may be able to gather data in a way that can detect sensor drift, but this is beyond the scope of this research.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors.
\item Use posterior sample to simulate a new smoother from each field.
\item Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this ``meta-smoother.''
\item Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
\end{enumerate}

This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.

@@

Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table 1, row 6).
Distances from boundaries were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).

Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 t/ha at the field edge and leveling off at about 3 t/ha (Figure 4 top row).
Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
Yield variability (Figure 4 bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.

@@

Pea crops were less well-represented in our data (N = 39 fields), and therefore the effects of boundaries tended to be less consistent.
Average pea yields (Figure 6, top row) were roughly 4.25 t/ha at the field boundary and increased to just below 5 t/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
Conversely, yield variability (Figure 6, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.

@@

Within individual fields, both the average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure 1).
There was a fairly large (0.75 t/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
This indicates that shelterbelts may have a positive effect on wheat yields (and possibly other crops not considered in this study), and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.

@@

Field boundary characteristics, such as vegetation height or species composition, may play an important role in mediating spillover, but our definitions of boundaries were necessarily loose, as we used satellite imagery to manually classify them.
Therefore, it may be that other aspects of field boundaries are more important for spillover of ecosystems services into fields.
For example, larger or more functionally diverse hedgerows tend to support more biodiversity than smaller hedgerows (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011; Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020), and some studies have shown that leaf litter and soil compaction influences predatory beetle communities (Sutter \emph{et al.} 2018).
These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
Finally, the relevant spatial scale of boundary effects are often related to the height or windward orientation of the shelterbelt (Kort 1988; Kowalchuk \& de Jong 1995), and precision yield data has the potential to show this in far greater detail than previous studies.

@@

We thank Trent Clark, Dean Hubbard, Alvin French, and Kristina Polziehn for providing yield data for us, and providing insight on their yield data.
We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
Finally, we thank Ben Woodcock and an anonymous second reviewer for their comments on the original manuscript.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-AAFC_data}{}%
Agriculture and Agri-Food Canada. (2022). \emph{{Annual Crop Inventory}}. Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}. Accessed 28 February 2022.

@@

\leavevmode\hypertarget{ref-pywell2015}{}%
Pywell, R.F., Heard, M.S., Woodcock, B.A., Hinsley, S., Ridding, L., Nowakowski, M. \& Bullock, J.M. (2015). Wildlife-friendly farming increases crop yield: evidence for ecological intensification. \emph{Proceedings of the Royal Society B: Biological Sciences}, 282, 20151740.

@@

### 2022-03-04 Manuscript/supplemental.Rmd
Updated manuscript

# Appendix A: R Code for Data Filtering {-}

@@

# Appendix B: R Code for Modelling {-}

>>>>

# Appendix A: Supplemental Information {-}

\beginsupplement

```{r fieldLocations, echo=FALSE, fig.cap='Locations of fields that data were collected from', out.width='100%'}  
  include_graphics('../Figures/Field Locations.png',dpi=NA)
```

```{r boundaryTypes, echo=FALSE, fig.cap='Boxplot of field boundary compositions across sampled fields. The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.', out.width='100%'}

bLengths %>% 
  mutate(croptype=datSourceFinal$crop) %>% 
  mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern Alberta','Central Alberta')) %>%
  group_by(location) %>% mutate(nFields=length(unique(filename))) %>% ungroup() %>% 
  mutate(location=paste0(location,' (N = ',nFields,')')) %>%
  select(filename,croptype,location,contains('prop_')) %>% 
  pivot_longer(contains('prop_')) %>% 
  mutate(name=gsub('prop_','',name),value=value*100) %>% na.omit() %>%
  mutate(name=factor(name,labels=c('Bare','Grassland','Other\nCrop','Shelterbelt','Standard','Wetland'))) %>% 
  group_by(croptype) %>% mutate(croptype=paste0(croptype,'\n(N = ',length(unique(filename)),')')) %>% 
  ggplot(aes(x=name,y=value,fill=croptype))+geom_boxplot()+
  facet_wrap(~location,ncol=1) +
  labs(y='Percentage of field boundary',fill='Crop')+
  scale_fill_manual(values=c('grey50','grey80','white'))+
  theme(axis.title.x = element_blank(),legend.position = 'bottom')

# bLengthSummary %>% #Table version
#   kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)

# Appendix B: R Code for Data Filtering {-}

@@

# Appendix C: R Code for Modelling {-}
